---
title: "econometrics_rtraining"
author: "Emily Finto"
date: "11/2/2021"
output: html_document
---
```{r}
library_list <- c(
  "AER",
  "MASS",
  "dplyr",
  "readxl",
  "scales",
  "stargazer",
  "mvtnorm",
  "plm",
  "rddtools",
  "tidyr",
  "dynlm",
  "forecast",
  "quantmod",
  "urca",
  "orcutt",
  "nlme",
  "vars",
  "fGarch"
)

for(library in library_list){
  
  if(!require(library, character.only = TRUE)){
    install.packages(library, dependencies = TRUE, repos = "http://cran.us.r-project.org")
    require(library, character.only = TRUE )
    
  }
}

```

#Ch 2. Probability Theory
##2.1 Random Variables and Probability Distributions
```{r 2_1_RandomVariables_ProbabilityDistributions, echo=FALSE}
probability <- rep(1/6, 6)

plot(probability,
     xlab= "outcomes",
     main="Probability Distribution")

cum_probability <- cumsum(probability)

plot(cum_probability,
     xlab="outcomes",
     main="Cumulative Probability Distribution")

#example: tossing a coin and observing the Bernoulli distributed var. What is the probability that we observed Head five times (k=5) when tossing the coin ten times with prob=0.5?
dbinom(x=5,
       size=10,
       prob=0.5)

#compute P(4 <= k <= 7)
sum(dbinom(x=4:7, size=10, prob=0.5))
pbinom(size=10, prob=0.5, q=7) - pbinom(size=10, prob=0.5, q=3)

#In this example, there are 11 possible outcomes for k.
k <- 0:10
k
probability <- dbinom(x=k,
                      size=10,
                      prob=0.5)
plot(x=k,
     y=probability,
     main="Probability Distribution Function")

#plot cumulative probabilities
prob <- pbinom(q=k,
               size=10,
               prob=0.5)

plot(x=k,
     y=prob,
     main="Cumulative Distributive Function")

#Expected Value, Mean, and Variance
mean(1:6)

#sampling with replacement
set.seed(1)
sample(1:6, 3, replace=TRUE)
mean(sample(1:6, 1000, replace=TRUE))
var(1:6)

#!!!!!Example
f <- function(x) 3 / x^4
g <- function(x) x*f(x)
h <- function(x) x^2 * f(x)

#area under the density curve
area <- integrate(f, 
                 lower = 1, 
                 upper = Inf)$value
area 

#compute E(x)
EX <- integrate(g,
                lower=1,
                upper=Inf)$value
EX

#compute Var(x)
Varx <- integrate(h,
                  lower=1,
                  upper=Inf)$value - EX^2
Varx

#The Normal Distribution
#define the standard normal PDF as an R function
f <- function(x){
  1/(sqrt(2*pi))*exp(-0.5*x^2)
}

quants <- c(-1.96, 0, 1.96)
f(quants)
#compare to see if the function computes standard normal densities
f(quants)==dnorm(quants)

#Chi-Squared Distribution
#plot where df (M) = 3
curve(dchisq(x, df=3),
      xlim=c(0,10),
      ylim=c(0,1),
      col="blue",
      ylab="",
      main="pdf and cdf of Chi-Squared Distribution, M=3")
curve(pchisq(x, df=3),
      xlim=c(0,10),
      add=TRUE,
      col="red")
legend("topleft",
       c("PDF", "CDF"),
       col=c("blue","red"),
       lty=c(1,1))

#curve shape depends on df
curve(dchisq(x, df = 1), 
      xlim = c(0, 15), 
      xlab = "x", 
      ylab = "Density", 
      main = "Chi-Square Distributed Random Variables")

# add densities for M=2,...,7 to the plot using a 'for()' loop 
for (M in 2:7) {
  curve(dchisq(x, df = M),
        xlim = c(0, 15), 
        add = T, 
        col = M)
}
legend("topright", 
       as.character(1:7), 
       col = 1:7 , 
       lty = 1, 
       title = "D.F.")

#the T distribution
# plot the standard normal density
curve(dnorm(x), 
      xlim = c(-4, 4), 
      xlab = "x", 
      lty = 2, 
      ylab = "Density", 
      main = "Densities of t Distributions")

# plot the t density for M=2, M=4, and M=25
curve(dt(x, df = 2), 
      xlim = c(-4, 4), 
      col = 2, 
      add = T)
curve(dt(x, df = 4), 
      xlim = c(-4, 4), 
      col = 3, 
      add = T)
curve(dt(x, df = 25), 
      xlim = c(-4, 4), 
      col = 4, 
      add = T)
legend("topright", 
       c("N(0, 1)", "M=2", "M=4", "M=25"), 
       col = 1:4, 
       lty = c(2, 1, 1, 1))
#as df increases, the shape of the t distribution is closer to the normal bell curve

#F distribution
#find P(Y >= 2), for F distributed random var Y with number df=3 and denom df=14
pf(2, df1=3, df2=14, lower.tail=F)
#visualize this prob
x <- c(2, seq(2, 10, 0.01), 10)
y <- c(0, df(seq(2, 10, 0.01), 3, 14), 0)

curve(df(x, 3, 14),
     ylim=c(0, .8),
     xlim=c(0,10),
     ylab="Density",
     main="Density Function")
polygon(x, y, col="orange")
```

##2.2 Random Sampling and the Distribution of Sample Averages
```{r 2_2_RandomSampling_DistributionSampleAverages, echo=FALSE}
# Plot the density histogram
hist(sample.avgs, 
     ylim = c(0, 1.4), 
     col = "steelblue" , 
     freq = F, 
     breaks = 20)

# overlay the theoretical distribution of sample averages on top of the histogram
curve(dnorm(x, sd = 1/sqrt(n)), 
      col = "red", 
      lwd = "2", 
      add = T)

# number of repetitions
reps <- 10000

# set degrees of freedom of a chi-Square Distribution
DF <- 3 

# sample 10000 column vectors à 3 N(0,1) R.V.S
Z <- replicate(reps, rnorm(DF)) 

# column sums of squares
X <- colSums(Z^2)

# histogram of column sums of squares
hist(X, 
     freq = F, 
     col = "steelblue", 
     breaks = 40, 
     ylab = "Density", 
     main = "")

# add theoretical density
curve(dchisq(x, df = DF), 
      type = 'l', 
      lwd = 2, 
      col = "red", 
      add = T)

#Law of Large Numbers - Central Limit Theorem
#coin toss example
set.seed(1)
N <- 30000
Y <- sample(0:1, N, replace = T)

# Calculate R_n for 1:N
S <- cumsum(Y)
R <- S/(1:N)

# Plot the path.
plot(R, 
     ylim = c(0.3, 0.7), 
     type = "l", 
     col = "steelblue", 
     lwd = 2, 
     xlab = "n", 
     ylab = "R_n",
     main = "Converging Share of Heads in Repeated Coin Tossing")

# Add a dashed line for R_n = 0.5
lines(c(0, N), 
      c(0.5, 0.5), 
      col = "darkred", 
      lty = 2, 
      lwd = 1)
```

#Ch. 3 Review of Statistics
```{r 3_reviewstatistics, echo=FALSE}
#3.1 Estimations of the Population Mean
# plot the chi_12^2 distribution
curve(dchisq(x, df=12), 
      from = 0, 
      to = 40, 
      ylab = "density", 
      xlab = "hourly earnings in Euro")

# sample from the chi_12^2 distribution, use only the first observation
set.seed(1)
rsamp <- rchisq(n = 100, df = 12)
rsamp[1]

#is this a good estimator? You want unbiasedness, consistency, and variance/efficiency

####################
#3.3 Hypothesis Tests Concerning the Population Mean
##the p-value is the probability of drawing data and observing a corresponding test statistic that is at least as adverse to what is stated in the null
# plot the standard normal density on the interval [-4,4]
curve(dnorm(x),
      xlim = c(-4, 4),
      main = "Calculating a p-Value",
      yaxs = "i",
      xlab = "z",
      ylab = "",
      lwd = 2,
      axes = "F")

# add x-axis
axis(1, 
     at = c(-1.5, 0, 1.5), 
     padj = 0.75,
     labels = c(expression(-frac(bar(Y)^"act"~-~bar(mu)[Y,0], sigma[bar(Y)])),
                0,
                expression(frac(bar(Y)^"act"~-~bar(mu)[Y,0], sigma[bar(Y)]))))

# shade p-value/2 region in left tail
polygon(x = c(-6, seq(-6, -1.5, 0.01), -1.5),
        y = c(0, dnorm(seq(-6, -1.5, 0.01)),0), 
        col = "steelblue")

# shade p-value/2 region in right tail
polygon(x = c(1.5, seq(1.5, 6, 0.01), 6),
        y = c(0, dnorm(seq(1.5, 6, 0.01)), 0), 
        col = "steelblue")

##t stat
samplemean_act <- mean(
  sample(0:1, 
         prob = c(0.9, 0.1), 
         replace = T, 
         size = 100))
SE_samplemean <- sqrt(samplemean_act * (1 - samplemean_act) / 100)
mean_h0 <- 0.1

tstatistic <- (samplemean_act - mean_h0) / SE_samplemean
tstatistic

# prepare empty vector for t-statistics
tstatistics <- numeric(10000)

# set sample size
n <- 300

# simulate 10000 t-statistics
for (i in 1:10000) {
  
  s <- sample(0:1, 
              size = n,  
              prob = c(0.9, 0.1),
              replace = T)
  
  tstatistics[i] <- (mean(s)-0.1)/sqrt(var(s)/n)
  
}
# plot density and compare to N(0,1) density
plot(density(tstatistics),
     xlab = "t-statistic",
     main = "Estimated Distribution of the t-statistic when n=300",
     lwd = 2,
     xlim = c(-4, 4),
     col = "steelblue")

# N(0,1) density (dashed)
curve(dnorm(x), 
      add = T, 
      lty = 2, 
      lwd = 2)

##test for significance
# compute the p-value
pvalue <- 2 * pnorm(- abs(samplemean_act - mean_h0) / SE_samplemean)
pvalue < 0.05

# check critical value
qnorm(p=0.975)

# check whether the null is rejected using the t stat
abs(tstatistic) > 1.96

####################
#3.4 Confidence Intervals for the Population Mean
set.seed(1)
sampledata <- rnorm(100, 10, 10)
t.test(sampledata)$"conf.int"
t.test(sampledata)
#the sample mean is significantly different from 0 at the 5% level since My=0 is not an element of the 95% conf int and p-value is much less than 0.05 (we fail to reject the null)

####################
#3.5 Comparing Means from Different Populations
# draw data from two different populations with equal mean
set.seed(1)
sample_pop1 <- rnorm(100, 10, 10)
sample_pop2 <- rnorm(100, 10, 20)

# perform a two sample t-test
t.test(sample_pop1, sample_pop2)
#the two-sample t test does not reject the null, the difference in the mean of the populations is different from 0
```

#Ch. 4 Linear Regression with One Regressor
```{r 4_linearregression, echo=FALSE}
#4.1 Simple Linear Regression
STR <- c(15, 17, 19, 20, 22, 23.5, 25)
TestScore <- c(680, 640, 670, 660, 630, 660, 635)

plot(TestScore ~ STR)
abline(a=713, b=-3)

####################
#4.2 Estimating the Coefficients of the Linear Regression Model
data(CASchools)
CASchools$STR <- CASchools$students/CASchools$teachers
CASchools$score <- (CASchools$read + CASchools$math)/2
avg_STR <- mean(CASchools$STR) 
avg_score <- mean(CASchools$score)
sd_STR <- sd(CASchools$STR) 
sd_score <- sd(CASchools$score)
quantiles <- c(0.10, 0.25, 0.4, 0.5, 0.6, 0.75, 0.9)
quant_STR <- quantile(CASchools$STR, quantiles)
quant_score <- quantile(CASchools$score, quantiles)

# gather everything in a data.frame 
DistributionSummary <- data.frame(Average = c(avg_STR, avg_score), 
                                  StandardDeviation = c(sd_STR, sd_score), 
                                  quantile = rbind(quant_STR, quant_score))
DistributionSummary

plot(score~STR,
     data=CASchools,
     main="Scatterplot of TestScore and STR",
     xlab="STR (X)",
     ylab="Test Score (Y)")

cor(CASchools$STR, CASchools$score)

#OLS estimators
linear_model <- lm(score~STR, data=CASchools)
summary(linear_model)

plot(score~STR,
     data=CASchools,
     main="Scatterplor of TestScore and STR",
     xlab="STR (X)",
     ylab="Test Score (Y)",
     xlim=c(10,30),
     ylim=c(600,700))
abline(linear_model)

####################
#4.3 Measures of Fit
#R-squared measures how well the variance is explained by the model. The Standard Error of the Regression (SER) measures the standard deviation of the residuals.

####################
#4.4 Least Squares Assumptions
#1. The error term has conditional mean zero
#2. ... are independent and identically distributed draws from their joint distribution
#3. Large outliers are unlikely
```

#Ch. 5 Hypothesis Tests and Confidence Intervals in the Simple Linear Regression Model
```{r 5_TestsIntervals_SimpleLinearRegression, echo=FALSE}
#5.4 Heterosckedasticity and Homoskedasticity
# set up vector of x coordinates
set.seed(123)
x <- rep(c(10,15,20,25), each=25)
# initialize vector of errors
e <- c()

# sample 100 errors such that the variance increases with x
e[1:25] <- rnorm(25, sd = 10)
e[26:50] <- rnorm(25, sd = 15)
e[51:75] <- rnorm(25, sd = 20)
e[76:100] <- rnorm(25, sd = 25)

# set up y
y <- 720 - 3.3 * x + e

# Estimate the model 
mod <- lm(y ~ x)

# Plot the data
plot(x = x, 
     y = y, 
     main = "An Example of Heteroskedasticity",
     xlab = "Student-Teacher Ratio",
     ylab = "Test Score",
     cex = 0.5, 
     pch = 19, 
     xlim = c(8, 27), 
     ylim = c(600, 710))

# Add the regression line to the plot
abline(mod, col = "darkred")

# Add boxplots to the plot
boxplot(formula = y ~ x, 
        add = TRUE, 
        at = c(10, 15, 20, 25), 
        col = alpha("gray", 0.4), 
        border = "black")
#the variance of test scores increases as the student-teacher ratio increases
#Computation of Heteroskedasticity-Robust Standard Errors - set type to "HC1"

vcov <- vcovHC(linear_model, type="HC1")
vcov
#^gives variance-covariance matrix of coefficient estimates. We want the square root of the diagonal of the matrix, the standard error estimates

robust_se <- sqrt(diag(vcov))
robust_se

coeftest(linear_model, vcov.=vcov)

#5.5 The Gauss-Markov Theorem
#If key assumptions hold and errors are homoskedastic, then the OLS estimator is the best linear conditionally unbiased estimator (BLUE)
```

#Ch. 6 Regression Models with Multiple Regressors
```{r 6_multivariablemodels, echo=FALSE}
#6.4 OLS Assumptions in Multiple Regression
#1. Regressors are drawn such that the i.i.d. assumption holds
#2. The error term has a conditional mean = 0
#3. Large outliers are unlikely
#4. No perfect multicollinearity

```

#Ch. 7 Hypothesis Tests and Confidence Intervals in Multiple Regression
see textbook

#Ch. 8 Nonlinear Regression Functions
```{r 8_nonlinear, echo=FALSE}
#8.2 Nonlinear Functions of a Single Independent Variable
#You can make a polynomial function
linear_model<- lm(score ~ income, data = CASchools)

quadratic_model <- lm(score ~ income + I(income^2), data = CASchools)
coeftest(quadratic_model, vcov.=vcovHC, type="HC1")

plot(CASchools$income, CASchools$score,
     col="steelblue",
     pch=20,
     xlab="District Income (thousands of dollars)",
     ylab="Test Score",
     main="Estimated Linear and Quadratic Regression Functions")
abline(linear_model, col="black", lwd=2)
order_id <- order(CASchools$income)
lines(x=CASchools$income[order_id],
      y=fitted(quadratic_model)[order_id],
      col="red",
      lwd=2)

#or a log

#8.3 Interactions Between Independent Variables
#b/w two binary vars, a continuous and binary, or two continuous vars

```

#Ch. 9 Assessing Studies Based on Multiple Regression
Chapter 9 gave theoretical details on possible issues with regressions including:
- internal/external validity
- omitted variable bias
- misspecification of the functional form of the regression function
- measurement errors
- missing data and sample selection
- simultaneous causality

#Ch. 10 Regression with Panel Data
```{r 10_paneldata, echo=FALSE}
#10.3 Fixed Effects Regression
# Consider model where one variable (z) is time invariant but we still want to look at the change in other variables (x).
     # y=b0+b1x+b2z

# Have individual intercepts (ai, i=1,...,n) where each of these can be understood as the fixed effect of entity i. The variation in a comes from z.
    # a=b0+b2z
    # Model now (10.1): y=z+b1x+u
    
# The fixed effects regression model:
    # (10.3) Y=B1X+...+BkXk+a+u
    # where a are entity-specific intercepts that capture heterogeneities across entities.
  # Could also be written with dummy regressors:
    # (10.4) Y=B0+B1X+...+BkXk+(gamma)2D2+(gamma)3D3+...+(gamma)nDn+u
    
# Can also use entity demeaned data:
  # (mean)Y=B1(mean)X+(mean)a+(mean)u
  # *subtract from model 10.1:
  # Y-(mean)Y = B1(X-(mean)X)+(u-(mean)u)
  # (squiggle)Y = B1(squiggle)X+(squiggle)u

#Ex. relation between traffic fatality rates and beer taxes:
    # FatalityRate = B1BeerTax + StateFixedEffects + u

Fatalities$fatal_rate <- Fatalities$fatal / Fatalities$pop * 10000
fatal_fe_lm_mod <- lm(fatal_rate ~ beertax + state - 1, data = Fatalities)
fatal_fe_lm_mod

#apply OLS to the demeaned data
Fatalities_demeaned <- with(Fatalities,
            data.frame(fatal_rate = fatal_rate - ave(fatal_rate, state),
            beertax = beertax - ave(beertax, state)))
summary(lm(fatal_rate~beertax-1, data=Fatalities_demeaned))

#or use plm() which uses the entity-demeaned OLS algorithm, not the dummy approach

fatal_fe_mod <- plm(fatal_rate ~ beertax, 
                    data = Fatalities,
                    index = c("state", "year"), 
                    model = "within")

# print summary using robust standard errors
coeftest(fatal_fe_mod, vcov. = vcovHC, type = "HC1")

#10.4 Regression with Time Fixed Effects
# If there are only time fixed effects, the regression becomes:
    # Y = B0 + B1X + (delta)B2 + ... + (delta)BT + u
  # where T-1 dummies are included (B1 is omitted) since the model included an intercept.
#Time and Fixed effects (ie. some things are fixed across time while others are fixed across entities):
    # Y = B0 + B1X + (gamma)2D2 +...+ (gamma)nDT + (delta)2B2 +...+(delta)TBT + u

#Ex. FatalityRates = B1BeerTax + Stateeffects + TimeEffects + u

fatal_tefe_lm_mod <- lm(fatal_rate ~ beertax + state + year - 1, data = Fatalities)
fatal_tefe_lm_mod

fatal_tefe_mod <- plm(fatal_rate ~ beertax, 
                      data = Fatalities,
                      index = c("state", "year"), 
                      model = "within", 
                      effect = "twoways")

coeftest(fatal_tefe_mod, vcov = vcovHC, type = "HC1")

#the variables state and year are of the class factors, which the lm() function automatically converts into dummies. Excluding the intercept with "-1" means that lm() estimates the coefficients for n+(T-1) = 48-6=54 binary variables (6 years and 48 states). The plm() function only reports the coefficient on BeerTax.

#10.5 The Fixed Effects Regression Assumptions and Standard Errors for Fixed Effects Regression
#1. The error term has a conditional mean = 0
#2. The variables and errors are i.i.d. draws from their joint distribution
#3. Large outliers are unlikely
#4. There is no perfect multicollinearity

#Standard errors
#In the event of heteroskedasticity and autocorrelation, you need to use heteroskedasticity and autocorrelation-consistent (HAC) standard errors, and clustered standard errors are an example of this. Clusteres standard errors are crucial in empirical applications of fixed effect models. Use the coeftest() function with vcovHC() for the plm() model, as the coeftest() automatically computes clustered se for plm() but not lm()
class(fatal_tefe_lm_mod)
coeftest(fatal_tefe_lm_mod, vcov = vcovHC, type = "HC1")[1, ]

class(fatal_tefe_mod)
coeftest(fatal_tefe_mod, vcov = vcovHC, type = "HC1")

```

#Ch. 11 Regression with a Binary Dependent Variable
```{r 11_binarydepvar, echo=FALSE}
#11.2 Probit and Logit Regression
#linear probability assumes that the conditional probability function is linear


```





