---
title: "wrangle"
author: "Emily Finto"
date: "9/8/2021"
output: html_document
---

#Wrangle
How to clean and organize your data to prepare it for analysis

You will need the following packages:
```{r packages}

library(tidyverse)
library(nycflights13)
library(Lahman)
library(gapminder)
library(readr)
```

##10. Tibbles
```{r intro to tibbles}
as_tibble(iris)

tibble(
  x = 1:5,
  y = 1,
  z = x^2 + y
)

tb <- tibble(
  `:)` = "smile",
  ` ` = "space",
  `2000` = "number"
)
tb

tribble(
  ~x, ~y, ~z,
  #--|--|----
  "a", 2, 3.6,
  "b",1, 8.5
)

##subsetting
df <- tibble(
  x = runif(5),
  y = runif(5)
)
###Extract by name
df$x
df[["x"]]
###Extract by position
df[[1]]
###use them in a pipe
df %>% .$x
df %>% .[["x"]]

```

###Older Code
not all data works with tibbles
```{r older code}
class(as.data.frame(tb))
```

##11. Data Import
```{reading in data}
#library(readr)
heights <- read_csv("heights.txt")
view(heights)
```

parsing:
```{r parsing}
parse_integer(c("1","231",".","456"), na=".")

#changing the default of "." to "," for parsing numbers
parse_double("1.23")

parse_double("1,23", locale=locale(decimal_mark=","))

#parse factors, a vector of known levels
fruit <- c("apple","banana")
parse_factor(c("apple","banana","bananana"), levels=fruit)

challenge <- read_csv(
  readr_example("challenge.csv"),
  col_types = cols(
    x=col_double(),
    y=col_date()
  )
)
tail(challenge)
```

##12. Tidy data
These are just examples because I cannot access the example tables.

###Pivot example:
tidy4a <- table4a %>%
  pivot_longer(c(`1999`,`2000`), names_to="year", values_to="cases")
tidy4b <- table4b %>%
  pivot_longer(c(`1999`,`2000`), names_to="year", values_to="population")
left_join(tidy4a, tidy4b)

###Separate example:
table3 %>%
  separate(rate, into=c("cases", "population"), sep="/", convert=TRUE)

Separate defaults to separating the values at non-alphanumeric characters, but you can specify where you want the columns to break. Additionally, using convert=TRUE has R assess what type of value the new column is (ie. string, integer), instead of the default which keeps the new columns the same as the original.

table3 %>%
  separate(year, into=c("century", "year"), sep=2)

you can also separate based on what digit you want to separate. Positive numbers count from the beginning of the value, negative numbers start counting from the end.

###Unite
table5 %>%
  unite(new, century, year, sep="")

R defaults to uniting values with an underscore between them, use the sep="" command to get rid of the underscore.
```{r tidying case study}
#data from WHO
who <- read_csv("TB_provisional_notifications_2021-09-09.csv")

who1 <- who %>%
  pivot_longer(
    cols=m_01:q_4,
    names_to = "key",
    values_to = "cases",
    values_drop_na = TRUE
  )
who1

#realized this is different data, can't find the original...

````
##13. Relational data
looks at options for joins

##14. Strings
How to handle string variables like counting characters, combining, subsetting, and edit it generally (ie. changing it all to lowercase). You can specify the language as well with locale=[language abbreviation]. It also reviews how to find letters or words within a list of values.






